Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-13 ms.tex
Small edits to section 3.1

Our work is a re-analysis of the SDSS DR7 data that 
overcomes numerous shortcomings of previous
analyses. In particular, we use direct population of galaxies in a
cosmological simulation so that delicate issues present in analytic
modeling, such as scale-dependent halo bias, are treated exactly.  The
simulation we use is based on the latest Planck cosmological
parameters \citep{planck13}, updating previous work. 
Furthermore, many important
differences between this work and the previous analysis by
\citet{zehavi_etal11} result from the fact that the Monte Carlo Markov
Chains used in \citet{zehavi_etal11} were too small and had not
converged (Z. Zheng \& I. Zehavi, private communication).  Therefore,
{\em our analysis supersedes previous work even in the case of
standard halo occupation models}.

@@

Prior to discussing our results using models that include assembly bias, we present 
results of standard HOD analyses that include no model for assembly bias. The results of 
the standard HOD analyses and all other analyses are shown in the form of marginalized 
constraints on individual parameters in Table~\ref{table:parameters}. We compare our parameter 
constraints to the standard HOD analysis performed by \citet{zehavi_etal11} in Table~\ref{table:parameters} 
as well. An example of the inferred posteriors for the HOD parameters is shown in Figure~\ref{fig:Mr20triangle}. 
The left-hand panels of Figure~\ref{fig:Mr19samples}, Figure~\ref{fig:Mr20samples}, and Figure~\ref{fig:Mr21samples} 
show the projected correlation function data along with predictions for $\wprp(\rp)$ from 50 randomly-selected 
models from the MCMC chains. Note that the significant covariance in the data makes it difficult to 
determine the quality of fit from visual inspection of these figures.

@@

The inferred parameters from our standard analyses differ in several ways from the \citet{zehavi_etal11} 
analysis. Firstly, in our re-analysis of the projected clustering data, we generally find all mass scales to 
be slightly higher than in the work of \citet{zehavi_etal11}. This difference is largely due to the slightly 
different cosmologies adopted in this work. The most important difference 
are in the values of $\Omega_{\rm M}$, and $\sigma_8$.  
\citet{zehavi_etal11} assumed $\Omega_{\rm M}=0.25$ and $\sigma_8=0.8$, 
whereas in the present work, we use the BolshoiP simulation in which $\Omega_{\rm M}=0.307$ and $\sigma_8=0.82$. 
Slightly larger mass scales are necessary in an analysis with higher $\Omega_{\rm M}$ and $\sigma_8$ in order 
to maintain galaxy number densities with larger halo number densities.

@@

Additionally, we have confirmed with a subset of the authors of \citet{zehavi_etal11} 
that the number of MCMC samples they included in their analysis was 
insufficient in a number of cases and that this can lead to a significant underestimation of the uncertainties 
on the inferred parameters, especially $\log (M_{\rm min})$ and $\sigma_{\log M}$ 
(Z. Zheng \& I. Zehavi, private communication). The analysis of \citet{zehavi_etal11} 
used $10^4$ samples, whereas we find several $\times 10^6$ samples are often necessary for convergence. 
Additionally, we have recreated qualitatively similar behavior by considering only small subsets of our 
full MCMC chains. Consequently, insufficient sampling of the posterior seems to be the likely resolution 
of the discrepancies between our work and that of \citet{zehavi_etal11}.

@@

As is evident in Figure~\ref{fig:Mr20triangle}, $\log (M_{\rm min})$ and $\sigma_{\log M}$ 
share a relatively narrow degeneracy as well. This degeneracy is largely induced by the 
measured number density of the sample. Increasing $\log (M_{\rm min})$ decreases galaxy 
number density, but this can be compensated by an increase in $\sigma_{\log M}$, which 
places galaxies in a fraction of the considerably more numerous halos with masses less 
than $M_{\rm min}$. The consequence is that $\log (M_{\rm min})$ and $\sigma_{\log M}$ are 
degenerate with each other such that most of the posterior probability lies in a narrow band 
along which $\log (M_{\rm min})$ and $\sigma_{\log M}$ are positively correlated, as shown in 
Fig.~\ref{fig:Mr20triangle}. In the following plots, we suppress the parameter $\sigma_{\log M}$,  
in order to increase the clarity of the plots, because the viable range of $\sigma_{\log M}$ is 
determined by this simple degeneracy with $\log (M_{\rm min})$.

The results of this subsection demonstrate that we achieve reasonable fits to projected galaxy clustering 
data using direct HOD population of a high-resolution numerical simulation of structure formation. These results 
also update and supersede existing constraints in the literature in at least three respects. First, we work within the 
best-fit Planck cosmology. Second, we perform our parameter inference analysis using direct population of halos 
identified in a numerical simulation of cosmological structure formation (BolshioP). This greatly mitigates modeling 
uncertainties associated with nonlinear density field evolution, scale-dependent halo bias, halo exclusion, 
or other effects that have been difficult to incorporate into analytical halo models with high precision. Third, 
we have explored the posteriors of the parameters with significantly more samples (roughly two orders of magnitude), 
thereby mitigating errors on inferred parameters and their errors induced by insufficient sampling of the 
posterior.

>>>>

Our work is a re-analysis of the SDSS DR7 data that 
overcomes numerous shortcomings of previous
analyses. In particular, we use direct population of galaxies in a
cosmological simulation so that delicate issues present in analytic
modeling, such as scale-dependent halo bias, are treated exactly.  The
simulation we use is based on the latest Planck cosmological
parameters \citep{planck13}, updating previous work. 
Furthermore, many important
differences between this work and the previous analysis by
\citet{zehavi_etal11} result from the fact that the Monte Carlo Markov
Chains used in \citet{zehavi_etal11} were not converged 
(Z. Zheng, private communication). Therefore,
{\em our analysis supersedes previous work even in the case of
standard halo occupation models}.

@@

Prior to discussing our results using models that include assembly bias, we present 
results of standard HOD analyses that include no model for assembly bias. The results of 
the standard HOD analyses and all other analyses are shown in the form of marginalized 
constraints on individual parameters in Table~\ref{table:parameters}. We compare our parameter 
constraints to the standard HOD analysis performed by \citet{zehavi_etal11} in Table~\ref{table:parameters} 
as well. An example of the inferred posteriors for the HOD parameters is shown in Figure~\ref{fig:Mr20triangle}. 
The left-hand panels of Figure~\ref{fig:Mr19samples}, Figure~\ref{fig:Mr20samples}, and Figure~\ref{fig:Mr21samples} 
show the projected correlation function data along with predictions for $\wprp(\rp)$ from 50 randomly-selected 
models from the MCMC chains within $\Delta \chi^2 \le 1$ of the best-fit model. 
Note that the significant covariance in the data makes it difficult to 
determine the quality of fit from visual inspection of these figures.

@@

The inferred parameters from our standard analyses differ in several ways from the \citet{zehavi_etal11} 
analysis. Firstly, in our re-analysis of the projected clustering data, we generally find all mass scales to 
be slightly higher than in the work of \citet{zehavi_etal11}. This difference is primarily due to the slightly 
different cosmological model used in this work as compared to that of \citet{zehavi_etal11}. 
The most important differences are in the values of $\Omega_{\rm M}$, and $\sigma_8$.  
\citet{zehavi_etal11} assumed $\Omega_{\rm M}=0.25$ and $\sigma_8=0.8$, 
whereas in the present work, we use the BolshoiP simulation in which $\Omega_{\rm M}=0.307$ and $\sigma_8=0.82$. 
Slightly larger mass scales are necessary in an analysis with higher $\Omega_{\rm M}$ and $\sigma_8$ in order 
to maintain galaxy number densities fixed with larger halo number densities.

@@

Additionally, we have confirmed with a subset of the authors of \citet{zehavi_etal11} 
that the MCMC chains used in a number of their analyses may not have been 
properly converged and that this may have led to a significant underestimation of the 
uncertainties on the inferred parameters, especially $\log (M_0)$, $\log (M_{\rm min})$, 
and $\sigma_{\log M}$ (Z. Zheng, private communication). The analysis of \citet{zehavi_etal11} 
used $10^4$ samples, whereas we find several $\times 10^6$ samples are often necessary for convergence. 
The cause of the problem is likely a restrictive proposal distribution that causes the chain to 
diffuse through the posterior distribution only very slowly. We have recreated qualitatively similar 
behavior by considering only small subsets of our full MCMC chains. Consequently, insufficient 
sampling of the posterior seems to be the likely resolution 
of the discrepancies between our work and that of \citet{zehavi_etal11}.

@@

As is evident in Figure~\ref{fig:Mr20triangle}, $\log (M_{\rm min})$ and $\sigma_{\log M}$ 
share a relatively narrow degeneracy as well. This degeneracy is largely induced by the 
measured number density of the sample. Increasing $\log (M_{\rm min})$ decreases galaxy 
number density, but this can be compensated by an increase in $\sigma_{\log M}$, which 
places galaxies in a fraction of the considerably more numerous halos with masses less 
than $M_{\rm min}$. The consequence is that $\log (M_{\rm min})$ and $\sigma_{\log M}$ are 
degenerate with each other such that most of the posterior probability lies in a narrow band 
along which $\log (M_{\rm min})$ and $\sigma_{\log M}$ are positively correlated. 
In the following plots, we exclude the parameter $\sigma_{\log M}$,  
in order to increase the clarity of the plots, because the viable range of $\sigma_{\log M}$ is 
determined by this simple degeneracy with $\log (M_{\rm min})$.

The results of this subsection demonstrate that we achieve reasonable fits to projected galaxy clustering 
data using direct HOD population of a high-resolution numerical simulation of structure formation. These results 
also update and supersede existing constraints in the literature in at least three respects. First, we work within the 
best-fit Planck cosmology. Second, we perform our parameter inference analysis using direct population of halos 
identified in a numerical simulation of cosmological structure formation (BolshioP). This greatly mitigates modeling 
uncertainties associated with nonlinear density field evolution, scale-dependent halo bias, halo exclusion, 
or other effects that have been difficult to incorporate into analytical halo models with high precision. Third, 
we have explored the posteriors of the parameters with significantly more samples (roughly two orders of magnitude), 
thereby mitigating errors on inferred parameters caused by insufficient sampling of the 
posterior by \citet{zehavi_etal11}.
